Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset 

### 2024-06-22 code/eda-code/eda_my version.qmd
Part 3

---
title: "Exploratory Data Analysis script"
date: "2024-02-07"
output: html_document
author: Antonio Flores
---

@@

# Setup
Necessary libraries for this script.
```{r}
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
library(tidyr)
library(forcats)
```

@@

```{r}
data_location <- here::here("data","processed-data","processeddata1.rds")
mydata <- readRDS(data_location)
```

@@

Quick look at positive cases vs negative cases of outcome variable (y_termSubscribed)
```{r}
plot(mydata$y_termSubscribed)
```

@@

```{r}
p9 = mydata |>
  filter(y_termSubscribed == "yes") |>
  ggplot(aes(age))+
  geom_bar()+
  labs(title = "yes")

@@

p9
p10
```

@@

```{r}
p11 = mydata |>
  filter(y_termSubscribed == "yes") |>
  ggplot(aes(x=fct_infreq(job)))+
  geom_bar()+
  coord_flip()+
  labs(title = "Term Subscribed = yes")

@@

>>>>

---
title: "Exploratory Data Analysis script"
date: "2024-06-20"
output: html_document
author: Antonio Flores
---

@@

# Setup
Necessary libraries for this script.
```{r}
#| echo: false
#| output: false
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
library(tidyr)
library(forcats)
library(corrplot)
```

@@

```{r}
#If this doesn't initially work, try restarting R and then proceeding
data_location <- here::here("data","processed-data","processeddata1.rds")
mydata <- readRDS(data_location)
```

@@

@@

```{r}
p9 = mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(age))+
  geom_bar()+
  labs(title = "yes")

@@

p9
p10

#Comparing how age daa compares from those who subscribed to a term deposit and those
#that did not
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=age))+
  facet_wrap(~y_termSubscribed, nrow=2)

@@

```{r}
p11 = mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(x=fct_infreq(job)))+
  geom_bar()+
  coord_flip()+
  labs(title = "Term Subscribed = yes")

@@

Trying to explore some additional aspects of the data

This shows the relationship between age and balance, grouped by education
```{r}
ggplot(data=mydata)+
  geom_point(mapping =aes(x=age, y=balance))+
  facet_wrap(~education, nrow=2)

Next, we can assess how age relates the number of jobs, grouped by job type.
```{r}
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=age))+
  facet_wrap(~job, nrow=2)
```

Digging into the Duration variable, we can see some very extreme values affecting this variable
```{r}
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=duration))

boxplot(mydata$duration)

Taking a quick look at some of the binary variables in the dataset
```{r}
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=default))
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=housing))
ggplot(data=mydata)+
  geom_bar(mapping =aes(x=loan))
```

Taking a look at frequency counts across different variables of interest
```{r}
hist(numericvar$age)
hist(numericvar$balance)
hist(numericvar$duration)
hist(numericvar$campaign)
hist(numericvar$pdays)
hist(numericvar$previous)
```

Examining which days are most common for recording a positive outcome (subscribing to 
a term deposit). Looks like later in the month has higher frequencies.

```{r}
p12= mydata |>
  filter(y_termSubscribed == 1) |>
  ggplot(aes(day))+
  geom_bar()+
  labs(title = "yes")

p12
```

We've analyzed which job types had the MOST positive outcomes, but now we want to 
compare which job type had the greates PERCENT of positive outcomes.

```{r}
posjob = mydata %>% 
  filter(y_termSubscribed==1) %>% 
  count(y_termSubscribed, job, sort=TRUE)

jobtable = mydata %>% count(job)

pctjob = posjob %>% 
  left_join(jobtable, by ="job")

colnames(pctjob)[3] <- "numofYES"
colnames(pctjob)[4] <- "numofBOTH"

ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
  geom_bar(stat="identity")+
  coord_flip()

ggplot(pctjob, aes(x=reorder(job, (numofYES/numofBOTH)), y=(numofYES/numofBOTH)))+
  geom_bar(stat="identity")+
  coord_flip()

```
On the first figure, we can see that jobs with the most positive outcomes are management, technician, and blue-collar, but on the second chart we can see that the job-types with the
highest percentages are 'student', 'retired', and 'unemployed.' There appears to be a noticable different between the total number of positive cases, and the job types which are most commonly recording positive cases.

### 2024-06-22 code/processing-code/processingfile-v1 - my version.qmd
Part 3

```{r}
library(readxl) 
library(dplyr) 
library(tidyr) 
library(skimr) 
library(here) 
```

@@

We are mainly converting character variables into factors so we can do some categorical analysis.
I had initially included converting months to a numeric variable but realized that it was fine as a factor variable.
```{r}
d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(y = as.factor(y))

>>>>

```{r}
#| echo: false
#| output: false
library(readxl) 
library(dplyr) 
library(tidyr) 
library(skimr) 
library(here) 
```

@@

We are mainly converting character variables into factors so we can do some categorical analysis.
I had initially included converting months to a numeric variable but realized that it was fine as a factor variable.
```{r}
d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(day = as.factor(day)) |>
  mutate(y = ifelse(y == "yes", 1, 0))

### 2024-06-22 products/manuscript/Manuscript.qmd
Part 3

# Summary/Abstract
This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.

@@

## General Background Information
Providing businesses with a model that will allow prioritization of consumers could save time and allow for a more accurate use of resources.

@@

The research question I plan to address with my analysis is which features or combination of features are the best predictors of consumers making a deposit. The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns. I don't currently have specific predictors in mind to begin with, as I plan to test all of the available features while also looking at some combinations or concatenation of features.

@@

I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already. Then I will test for correlation between different features and proceed from there.

## Schematic of workflow

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
```

## Data aquisition
Data retrieved from UCI ML Repository

@@

## Statistical analysis

@@

@@

@@

>>>>

# Summary/Abstract
This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank. After the data has been cleaned and prepared for analysis, exploratory data analysis will be performed to gather more information regarding the shape, size, and behavior of different variables, to gauge their usefulness in a prediction model. Additionally, several different statistical tests will be implemented to identify which classes within different variables affect the outcome variable. Finally (_haven't started on this yet_), with some additional data preprocessing, this project will test different machine learning models and conclude with the resulting findings.

@@

## General Background Information
Providing businesses with a model that will allow prioritization of consumers and/or demographics has great potential in improving resource management and future marketing campaigns, as well as increasing efficient spending.

@@

The research question I plan to address with my analysis is: which features or combination of features are the best predictors of consumers making a deposit? The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns. Currently, I plan to investigate all variables, but I am specifically interested in both job type, education and age.

@@

Cleaning -> This includes converting numeric variables to factors, changing column names, and converting the positive outcome varible to 1 and 0.

Exploring -> Using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.

## Data aquisition
The dataset for this project was retrieved from UCI ML Repository in CSV form.
Additionally, I created a codebook based on data from the same source.

@@

@@

@fig-result4 shows barplot of age data for the full dataset

```{r}
#| label: fig-result4
#| fig-cap: "Age data for those with positive classifcation"
#| echo: FALSE
knitr::include_graphics(here("results","figures","yes-no-age.png"))
```

@fig-result3 shows barplot of age data for only those who subscribed to a term deposit

```{r}
#| label: fig-result3
#| fig-cap: "Age data for those with positive classifcation"
#| echo: FALSE
knitr::include_graphics(here("results","figures","yes-age.png"))
```

@fig-result5 shows barplot of most common days of the month to record a positive outcome.

```{r}
#| label: fig-result5
#| fig-cap: "Days of the month for positive outcomes"
#| echo: FALSE
knitr::include_graphics(here("results","figures","days-yes.png"))
```

@fig-result6 Job type for count of positive outcomes.

```{r}
#| label: fig-result6
#| fig-cap: "Job type for count of positive outcomes."
#| echo: FALSE
knitr::include_graphics(here("results","figures","jobs-count-yes.png"))
```

@fig-result7 Job type for percent of positive outcomes per job type.

```{r}
#| label: fig-result7
#| fig-cap: "Job type for percent of positive outcomes per job type."
#| echo: FALSE
knitr::include_graphics(here("results","figures","jobs-percent-yes.png"))
```

@@

#WORK IN PROGRESS

@@

_Pending Data Preprocessing and ML Modeling_
_I plan to use recipies and some of the tinymodels for this section_
